### 2017-02-08 CO750-ApproximationAlgorithmsTopics/assignments/assignment_01.tex
Typed up 6a

\subsection{}

@@

\item $d(u,v) \leq d^T(u,v)$ for all $u,v\in V$

@@

\item Choose randomly a permutation $\pi$ of $V$.

@@

@@

@@

\item Return tree $T$ where $V(T) = \{s \in D_i : \forall i = 0, \dots, \delta\}$ and $E(T) = \{s_is_{i-1}  \in D_{i}\times D_{i-1}: s_{i-1} \subseteq s_i, \forall i = 0,\dots,\delta\}$, and $d(s_is_{i-1}) = 2^i$.

@@

Note that the clusters of $D_0$ are singletons because the bound on the maximum distance is $2^\beta 2^0 \leq 1$. Hence the clusters of $D_0$ can be identified with nodes in $V$. 

@@

Suppose $u,v$ are in the same cluster at level $D_i$ and are separated at level $D_{i-1}$. Let $w$ be the node identified with $u$ and $v$'s cluster $S_w$ at level $D_i$. Then $d(u,v) \leq d(u,w) + d(w,v) \leq 2^\beta 2^{i-1} + 2^\beta 2^{i-1} \leq 2^{i+1}$. On the other hand $d^T(u, S_w) \geq 2^i$ and $d^T(S_w, v) \geq 2^i$ so $d^T(u,v) \geq 2^i + 2^i = 2^{i+1}$. This proves part $1.$ of the theorem.

@@

&= 2^i + 2^i +  d^T(S^u_{i-1}, u) + d^T(S^v_{i-1}, v) \\

@@

Now we way that node $c$ {\it settles} $(u,v)$ at level $D_i$ if $c$ is the first node with respect to $\pi$ to which one of $u$ or $v$ gets assigned at level $D_i$. We say that node $c$ {\it cuts} $(u,v)$ at level $D_i$ if exactly one of $u,v$ is at distance at most $2^\beta 2^i$ from $c$. If $u,v$ are separated at level $D_i$ then there exists some $c$ which both settles and cuts $u,v$ at level $D_i$.

@@

Note that $$E[d^T(u,v)] \leq \sum_{i=0}^{\delta - 1} \sum_{c \in V} Pr [ \text{$c$ settles (A) and cuts (B) $u,v$ at $D_i$}] 2^{i+2}.$$ The bound simply follows from our observation, and the prior definitions. Recall that $Pr[A \cap B] = P[A\mid B] Pr[B]$.

@@

We first analyze $Pr[B]$. Without loss we may assume $d(c,u) \leq d(c,v)$. For $c$ to cut $u,v$ we need $d(c,u) \leq 2^\beta 2^{i-1} < d(c,v)$ by definition. Thus
$$Pr[B] = \frac{| [2^{i-1},2^i] \cap [d(c,u),d(c,v)]|}{|[2^{i-1},2^i]|} = \frac{| [2^{i-1},2^i] \cap [d(c,u),d(c,v)]|}{2^{i-1}}$$.
Now we need $Pr[A \mid B]$. Let us order the nodes of $V$ according to their distance to $u,v$. That is, $V = \{c_1, \dots, c_n\}$ where $\min\{d(c_i, v), d(c_i, u)\} \leq \min \{d(c_{i+1}, v), d(c_{i+1}, u)\}$ for all $i$. If $c= c_k$ then the probability that $c$ comes before $c_1, \dots, c_{k-1}$ in $\pi$ is at most $\frac{1}{k}$. In order for $c$ (given that $c$ cuts $u,v$) to settle $u,v$ at least that has to happen. Hence

@@

There exists a tree $T' = (V,F)$ such that $d^T(u,v) \leq d^{T'}(u,v) \leq 4 d^T(u,v).$
Pick $v \in V$ with a parent $w \not \in V$ and contract edge $vw$ in $T$. Repeat until all nodes of $T$ are in $V$. Multiply by $4$ the weights on all remaining edges. Call the resulting tree $T'$. Note that clearly $d^{T'}(u,v) \leq 4d^T(u,v)$ as contracting only shortens the distances. Furthermore, in the worst case $d^{T'}(u,v) = 2^i \cdot 4$ if contraction results in contracting $u$ and $v$ right up to the cluster where they were separated (that is, contracting preserves the largest edge distance on the $uv$-path in $T$). But 
$$2^i \cdot 4 = 2^i + 2^i + 2^i + 2^i \geq d(S^u_{i-1}, S_i) + d(S^v_{i-1}, S_i) +  d(S^u_{i-1}, u) + d(S^v_{i-1}, v) = d^T(u,v).$$
Hence the theorem holds.
\end{document}

>>>>

@@

\item $d(u,v) \leq d^T(u,v)$ for all $u,v\in W$

@@

\item Choose randomly a permutation $\pi$ of $W$.

@@

\item For each node such that its distance from each vertex in $W$ is greater that $2^\beta \cdot 2^{i-1}$ assign it its own cluster.

@@

\item Add a node to $T$ for each new cluster in $D_i$ and add edges to $T$ from each such node to the parent node in $T$ corresponding to their common cluster in $D_{i+1}$.
\item For the clusters formed in Step $(b)$ assign distance $0$ to the new edge in $T$. For all others (ie those from $(a)$) assign distance $2^{i+1}$.

@@

\item Return tree $T$.

@@

Note that the clusters of $D_0$ are singletons because the bound on the maximum distance is $2^\beta 2^0 \leq 1$. Hence all nodes of $V$ are invariably separated at some level by the end of operation of the algorithm.

@@

Suppose $u,v \in W$ are in the same cluster at level $D_i$ and are separated at level $D_{i-1}$. Let $w$ be the node identified with $u$ and $v$'s cluster $S_w$ at level $D_i$. Then
$$d(u,v) \leq d(u,w) + d(w,v) \leq 2^\beta 2^{i-1} + 2^\beta 2^{i-1} \leq 2^{i+1}.$$ The first inequality follows from the triangle inequality. The second inequality follows since the distance from $u$ (similarly $v$) to the next node on the $uw$-path is $2^{i-1}$ and the distance of the rest of the path is at most $\sum_{k=0}^{i-2} 2^k \leq 2^{i-1}$. On the other hand $d^T(u, S_w) \geq 2^i$ and $d^T(S_w, v) \geq 2^i$ so $d^T(u,v) = d^T(u, S_w) + d^T(S_w, v) \geq 2^i + 2^i = 2^{i+1}$. This proves part $1.$ of the theorem.

@@

&\leq \max\{2^i,0\} + \max\{2^i,0\} +  d^T(S^u_{i-1}, u) + d^T(S^v_{i-1}, v) \\

@@

Now we say that node $c$ {\it settles} $(u,v)$ at level $D_i$ if $c$ is the first node with respect to $\pi$ to which one of $u$ or $v$ gets assigned at level $D_i$. We say that node $c$ {\it cuts} $(u,v)$ at level $D_i$ if exactly one of $u,v$ is at distance at most $2^\beta 2^i$ from $c$. If $u,v$ are separated at level $D_i$ then either there exists some $c \in W$ which both settles and cuts $u,v$ at level $D_i$ (in which case they were separated by Step $4(a)$), or every $c \in W$ does not settle $u,v$ (in which case they were separated by Step $4(b)$).

@@

Note that in the first case the distance between $u$ and $v$ is at most $2^{i+2}$ by our previous observation. In the second case both $u$ and $v$ were assigned their own cluster in $D_i$ and given distance $0$ to their common ancestor. Hence $d^T(u,v) = 0$ in this second case. So we have the following bound on $E[d^T(u,v)]$:
 \begin{align*}E[d^T(u,v)] &\leq \sum_{i=0}^{\delta - 1} (\sum_{c \in W} Pr [ \text{$c$ settles $u,v$ and  $c$ cuts $u,v$ at $D_i$}] 2^{i+2} + Pr[\text{all $c \in W$ do not settle $u,v$}]\cdot 0) \\&= \sum_{i=0}^{\delta - 1} \sum_{c \in W} Pr [ \text{$c$ settles $u,v$ and  $c$ cuts $u,v$ at $D_i$}] 2^{i+2}.
 For $c \in W$, $u,v \in V$, and level $D_i$,  let $A$ be the event that $c$ settles $u,v$ and let $B$ be the event that $c$ cuts $u,v$ at $D_i$. We are interested in $Pr[A \cap B]$, or equivalently, $Pr[A \mid B] \cdot Pr[B]$.

@@

We first analyze $Pr[B]$. Without loss we may assume $d(c,u) \leq d(c,v)$. For $c$ to cut $u,v$ we need $d(c,u) \leq 2^\beta 2^{i-1} < d(c,v)$ by definition.  Since we sample $2^\beta$ uniformly at random, the probability of this happening is simply $\frac{1}{2^{i-1}}$ times the length of the interval intersecting $[2^{i-1},2^i]$ with $[d(c,u), d(c,v)]$. More formally:
$$Pr[B] = \frac{| [2^{i-1},2^i] \cap [d(c,u),d(c,v)]|}{|[2^{i-1},2^i]|} = \frac{| [2^{i-1},2^i] \cap [d(c,u),d(c,v)]|}{2^{i-1}}.$$
Now we need $Pr[A \mid B]$. Let us order the nodes of $W$ according to their distance to $u,v$. That is, $W = \{c_1, \dots, c_n\}$ where $\min\{d(c_i, v), d(c_i, u)\} \leq \min \{d(c_{i+1}, v), d(c_{i+1}, u)\}$ for all $i$. If $c= c_k$ then the probability that $c$ comes before $c_1, \dots, c_{k-1}$ in $\pi$ is at most $\frac{1}{k}$. In order for $c$ (given that $c$ cuts $u,v$) to settle $u,v$ at least that has to happen. Hence

@@

%Pb
\subsection{b}
\begin{theorem}
There exists a tree $T' = (V,F)$ such that $d^T(u,v) \leq d^{T'}(u,v) \leq 4 d^T(u,v).$
\end{theorem}
\begin{proof}
Pick $v \in V$ with a parent $w \not \in V$ and contract edge $vw$ in $T$. Repeat until all nodes of $T$ are in $V$. Multiply by $4$ the weights on all remaining edges. Call the resulting tree $T'$. Note that clearly $d^{T'}(u,v) \leq 4d^T(u,v)$ as contracting only shortens the distances. Furthermore, in the worst case $d^{T'}(u,v) = 2^i \cdot 4$ if contraction results in contracting $u$ and $v$ right up to the cluster where they were separated (that is, contracting preserves the largest edge distance on the $uv$-path in $T$). But 
$$2^i \cdot 4 = 2^i + 2^i + 2^i + 2^i \geq d(S^u_{i-1}, S_i) + d(S^v_{i-1}, S_i) +  d(S^u_{i-1}, u) + d(S^v_{i-1}, v) = d^T(u,v).$$
Hence the theorem holds.
\end{proof}
\end{document}
